---
title: "Assignment2_AW"
author: "aw"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

For this assignment I collected a small portion of data on popular movies. I stored the data in a locally hosted mySQL database and experimented with two different ways of imputing missing data (median and mean) to see how it might impact bias or skew analysis. Below, I will walk you through the data collection method, the database setup, and the r functions used to sort, impute, and compare the two different methods.

## Data Collection

You can access the original survey here: \<<https://forms.gle/LE8KtSGwFtNzi9cp8>\>. Information gathered included general demographic info (age-range, gender, movie_genre interests) as well as each respondents ratings of 6 different movies released in the last year. Respondents had the option of leaving the rating question blank if they hadn't seen the movie.

## Database Setup

Survey answers were compiled in a csv file which you can find here: \<<https://github.com/awrubes/Assignment2_Data607>\> "movieratingresponses.csv". You will need to download the mySQL script "createmySQLDatabase.sql". Make sure to change the file path to match where you are storing your csv file. Once executed, you should have a database called 'Movie Ratings' with three distinct tables: 'movie_ratings', 'respondents' and 'survey_responses'.

## R Setup

In order to access the mySQL database, you will need to install the following packages:

```{r libs, results = "hide"}
library(DBI)
library(dplyr)
library(RMySQL)
Sys.getenv()
```

You'll need to provide the user name and password for the root user (or whichever user has proper privileges) and connect to the database.

```{r dbsetup, results = "hide"}
user_name <- Sys.getenv("MYSQL_USER")
pass_word <- Sys.getenv("MYSQL_PASS")

conn <- dbConnect(RMySQL::MySQL(), dbname = "Movie Ratings", host = "localhost", user = user_name, password = pass_word)

dbListTables(conn)
```

## Creating a Dataframe

Since rating and respondent data was stored in two separate tables that reference each other using respondent unique ID's, you'll need to combine the tables.

```{r query}
query <- "SELECT r.name, r.gender, r.age_range, r.favorite_genres, m.movie_name, m.rating
          FROM respondents r
          JOIN movie_ratings m
          ON r.respondent_id = m.respondent_id;"
data <- dbGetQuery(conn, query)

head(data, 10)
```

```{r missing}
missing_ratings <- sum(is.na(data$rating))
print(missing_ratings)
```

There are 9 missing ratings from the data. Next we'll use median and mean to imputation. First, it's important to understand that mean/median imputation is useful when the data is MCAR (missing completely at random), and since a person's decision to see a movie or not is usually not a random one, using this means of imputation may not be the most accurate.

## Imputing Missing Data

```{r calculate}
data_mean_imputed <- data %>%
  group_by(movie_name) %>%
  mutate(rating = ifelse(is.na(rating), mean(rating, na.rm = TRUE), rating))

data_median_imputed <- data %>%
  group_by(movie_name) %>%
  mutate(rating = ifelse(is.na(rating), median(rating, na.rm = TRUE), rating))

head(data_mean_imputed)
head(data_median_imputed)

summary(data_mean_imputed$rating)
summary(data_median_imputed$rating)
```

``` r
dbDisconnect(conn)
```
